Viral Sequence Sorting Tools Evaluation

### 2022-09-27 EnvironmentalData/DrinkingWaterMetagenomes/identifying_viral_contigs_environmental_dw.Rmd
New visualizations of drinking water viruses.

```{r inputs}
# dataset name for file organization and outputs for metrics file
dataset_name <- "drinking_water"
# checkV
checkV_path <- "DrinkingWaterToolOutput/merged_checkV_allcontigs.tsv"
# VIBRANT
vibrant_path <- "DrinkingWaterToolOutput/vibrant_merged_contig_list_all.txt"
# DeepVirFinder
dvf_path <- "DrinkingWaterToolOutput/dvf_outputs_merged.tsv"
# Virsorter
vs_path <- "DrinkingWaterToolOutput/vs_merged_all.txt"
# Virsorter2
vs2_path <- "DrinkingWaterToolOutput/vs2_merged_file.txt"
# Kaiju
kj_path <- "DrinkingWaterToolOutput/kaju_outputs_merged.tsv"
# KB cutoff
KB_CUTOFF <- 3000

@@

## Merging
to do: this step isn't working properly
```{r}
viruses <- full_join(x=vs2_c, y=checkV, by = c("uniq_contig", "assembly"))
viruses <- full_join(x=viruses, y=dvf_c, by = c("uniq_contig", "assembly"))
viruses <- full_join(x=viruses, y=vb_c, by = c("uniq_contig", "assembly"))
viruses <- full_join(x=viruses, y=vs_c, by = c("uniq_contig", "assembly"))
viruses <- full_join(x=viruses, y=kj_c, by = c("uniq_contig", "assembly"))
```

@@

```{r}
viruses_filename <- 'merged_viral_id_outputs.csv'
write.csv(viruses, paste(dataset_name, viruses_filename, sep='_'))
```

@@

>>>>

```{r inputs}
# dataset name for file organization and outputs for metrics file
dataset_name <- "drinking_water"
# checkV
checkV_path <- "DrinkingWaterToolOutput/merged_checkV_allcontigs.tsv"
# VIBRANT
vibrant_path <- "DrinkingWaterToolOutput/vibrant_merged_contig_list_all.txt"
# DeepVirFinder
dvf_path <- "DrinkingWaterToolOutput/dvf_outputs_merged.tsv"
# Virsorter
vs_path <- "DrinkingWaterToolOutput/vs_merged_all.txt"
# Virsorter2
vs2_path <- "DrinkingWaterToolOutput/vs2_merged_file.txt"
# Kaiju
kj_path <- "DrinkingWaterToolOutput/kaiju_outputs_merged.tsv"
# KB cutoff
KB_CUTOFF <- 3000

@@

## Merging
```{r}
viruses <- full_join(x=vs2_c, y=checkV, by = c("uniq_contig", "assembly"))
viruses <- full_join(x=viruses, y=dvf_c, by = c("uniq_contig", "assembly"))
viruses <- full_join(x=viruses, y=vb_c, by = c("uniq_contig", "assembly"))
viruses <- full_join(x=viruses, y=vs_c, by = c("uniq_contig", "assembly"))
viruses <- full_join(x=viruses, y=kj_c, by = c("uniq_contig", "assembly"))
```

@@

```{r}
viruses_filename <- 'merged_viral_id_outputs.csv'
#write.csv(viruses, paste(dataset_name, viruses_filename, sep='_'))
```

@@

# Comparing all of the tool methods

```{r}
combos_list <- data.frame(toolcombo=rep(0, 64),
                          CheckV=rep(0, 64),
                          DVF=rep(0, 64),
                          Kaiju=rep(0, 64),
                          VIBRANT=rep(0, 64),
                          VS=rep(0, 64),
                          VS2=rep(0, 64))
p <- 1

for (i in c(0,1)){
  for (j in c(0,1)){
    for (k in c(0,1)){
      for (l in c(0,1)){
        for (m in c(0,1)){
          for (n in c(0,1)){
            combos_list$toolcombo[p] <- paste(i,j,k,l,m,n)
            combos_list$CheckV[p] <- i
            combos_list$DVF[p] <- j
            combos_list$Kaiju[p] <- k
            combos_list$VIBRANT[p] <- l
            combos_list$VS[p] <- m
            combos_list$VS2[p] <- n
            p <- p+1
          }
        }
      }
    }
  }
}

combos_list <- combos_list[-1,]
```

```{r}
viral_scores <- matrix(data=0, nrow=nrow(viruses), ncol=nrow(combos_list))
num_viruses <- data.frame(toolcombo=rep(0, nrow(combos_list)),
                          num_viruses=rep(0, nrow(combos_list)))

for (i in 1:nrow(combos_list)) {
  viral_scores[,i] <- getting_viral_set_1(viruses, include_vibrant = combos_list$VIBRANT[i],
                                            include_virsorter = combos_list$VS[i],
                                            include_virsorter2 = combos_list$VS2[i],
                                            include_tuning = combos_list$CheckV[i],
                                            include_kaiju = combos_list$Kaiju[i],
                                            include_deepvirfinder = combos_list$DVF[i])

  num_viruses$num_viruses[i] <- table(viral_scores[,i]>=1)[[2]]

  num_viruses$toolcombo[i] <- combos_list$toolcombo[i]
}

num_viruses$numtools <- str_count(num_viruses$toolcombo, "1")
num_viruses <- num_viruses[order(num_viruses$num_viruses, decreasing=F),]
num_viruses$toolcombo <- factor(num_viruses$toolcombo, levels = unique(num_viruses$toolcombo))
num_viruses$numtools <- as.factor(num_viruses$numtools)
```

```{r}
pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
ggplot(num_viruses, aes(x=toolcombo, y=num_viruses, 
                                  color=numtools, fill=numtools)) +
  geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
  ylab("Num Viruses Predicted")
```

```{r}
viral_scores_nozeros <- viral_scores[rowSums(viral_scores)>0,]
viral_scores_nozeros <- viral_scores_nozeros + 1
viral_scores_nozeros <- as.data.frame(viral_scores_nozeros)

colnames(viral_scores_nozeros) <- num_viruses$toolcombo
```

```{r}
library(phyloseq)
```

```{r}
tooldata <- num_viruses
rownames(tooldata) <- tooldata$toolcombo
```

```{r}
physeq_pooled <- phyloseq(otu_table(viral_scores_nozeros, taxa_are_rows = T),
                                     sample_data(tooldata))

ordination <- phyloseq::ordinate(physeq =physeq_pooled, method = "PCoA", distance = "bray")
phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
                          shape="numtools", color="num_viruses") + 
  geom_point(size = 3) +
  theme_bw() +
  geom_label(label=tooldata$toolcombo)

phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
                          shape="numtools", color="num_viruses") + 
  geom_point(size = 3) +
  theme_bw()
```
to do: try coloring above based on the F1 scores of the testing set on each combination

```{r}
write_tsv(num_viruses, "dw_num_viruses_by_tool_combo.tsv")
num_viruses <- read_tsv("dw_num_viruses_by_tool_combo.tsv")
```

```{r}
accuracy_scores <- read_tsv("../../TestingPipeline/Scripts/20220927_accuracy_scores.tsv")
```

```{r}
num_viruses_comp <- full_join(num_viruses, accuracy_scores[accuracy_scores$testing_set_index==1,], by=c("toolcombo", "numtools"))
```

```{r}
num_viruses_comp$numtools <- as.factor(num_viruses_comp$numtools)
```

```{r}
ggplot(num_viruses_comp, aes(x=num_viruses, y=F1, 
                                  color=numtools, fill=numtools)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  ylab("F1") +
  xlab("Num Viruses Predicted")

ggplot(num_viruses_comp, aes(x=num_viruses, y=precision, 
                                  color=numtools, fill=numtools)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  ylab("Precision") +
  xlab("Num Viruses Predicted")

ggplot(num_viruses_comp, aes(x=num_viruses, y=recall, 
                                  color=numtools, fill=numtools)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  ylab("Recall") +
  xlab("Num Viruses Predicted")
```

### 2022-09-27 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
New visualizations of drinking water viruses.

combination of tools list
```{r}
combos_list <- read_csv("combinations_list.csv", col_names = T)
```

@@

## Calculate the performance of each pipeline
```{r}
accuracy_scores <- build_score_list(viruses[viruses$Index==1], combos_list)
accuracy_scores <- data.frame(testing_set_index=rep(0, nrow(combos_list)*10),
                      precision=rep(0, nrow(combos_list)*10),
                       recall=rep(0, nrow(combos_list)*10),
                       F1=rep(0, nrow(combos_list)*10),
                       MCC=rep(0, nrow(combos_list)*10), 
                      AUC=rep(0, nrow(combos_list)*10),
                      fungal_FP=rep(0, nrow(combos_list)*10),
                      bacterial_FP=rep(0, nrow(combos_list)*10),
                      viral_FN=rep(0, nrow(combos_list)*10))
accuracy_scores <- cbind(testing_set_index=rep(1, nrow(combos_list)),
                              build_score_list(viruses[viruses$Index==1], combos_list))
for (i in 2:10) {
  accuracy_scores <- rbind(accuracy_scores,
                           cbind(testing_set_index=rep(i, nrow(combos_list)),
                              build_score_list(viruses[viruses$Index==i], combos_list)))
}
```

>>>>

combination of tools list
```{r}
combos_list <- data.frame(toolcombo=rep(0, 64),
                          CheckV=rep(0, 64),
                          DVF=rep(0, 64),
                          Kaiju=rep(0, 64),
                          VIBRANT=rep(0, 64),
                          VS=rep(0, 64),
                          VS2=rep(0, 64))
p <- 1

for (i in c(0,1)){
  for (j in c(0,1)){
    for (k in c(0,1)){
      for (l in c(0,1)){
        for (m in c(0,1)){
          for (n in c(0,1)){
            combos_list$toolcombo[p] <- paste(i,j,k,l,m,n)
            combos_list$CheckV[p] <- i
            combos_list$DVF[p] <- j
            combos_list$Kaiju[p] <- k
            combos_list$VIBRANT[p] <- l
            combos_list$VS[p] <- m
            combos_list$VS2[p] <- n
            p <- p+1
          }
        }
      }
    }
  }
}
```

@@

## Calculate the performance of each pipeline
```{r}
accuracy_scores <- data.frame(testing_set_index=rep(0, nrow(combos_list)*10),
                      precision=rep(0, nrow(combos_list)*10),
                       recall=rep(0, nrow(combos_list)*10),
                       F1=rep(0, nrow(combos_list)*10),
                       MCC=rep(0, nrow(combos_list)*10), 
                      AUC=rep(0, nrow(combos_list)*10),
                      fungal_FP=rep(0, nrow(combos_list)*10),
                      bacterial_FP=rep(0, nrow(combos_list)*10),
                      viral_FN=rep(0, nrow(combos_list)*10))
accuracy_scores <- cbind(testing_set_index=rep(1, nrow(combos_list)),
                              build_score_list(viruses[viruses$Index==1], combos_list))
for (i in 2:10) {
  accuracy_scores <- rbind(accuracy_scores,
                           cbind(testing_set_index=rep(i, nrow(combos_list)),
                              build_score_list(viruses[viruses$Index==i], combos_list)))
}
```
